\documentclass[a4]{book}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{minted}
\usepackage{natbib}

\bibliographystyle{unsrtnat}

\author{Richard Neher\\Biozentrum, University of Basel}
\title{Physics of Life I}

\graphicspath{{figures/}}

\begin{document}
\maketitle

\chapter*{Preamble}

The laws of physics govern living matter as much as they govern inanimate matter.
However, it is often hard to see how physical principles enter in biology because living matter is constantly in flux, heterogeneous, and never at equilibrium.
Nevertheless, physics can elucidate many seemingly complex phenomena in biology.

In this course -- co-taught with Sebastian Hiller -- we will attempt to how show how biology can not only be described in terms of ``parts lists'' and ``cartoon''-like sketches, but understood in terms of quantitative and predictive relationships.
Such quantitative methods are becoming more and more important in modern biology and require a solid mathematical toolbox and often programming skills.
I encourage you to take a close look again at:
\begin{itemize}
	\item Mathematics: your high-school math and ``Mathematische Methoden I\&II''
	\item Physics: Mechanics, fluid dynamics, and thermodynamics from ``Physics I''
	\item Programming: python note-books and basic programming as you learned in programming and computational courses in the first year.
\end{itemize}

{\bf Lecture notes:}
These lecture notes were first written a few years ago and are adapted to the content of the current course to some extent.
I might make some changes to them over time, add additional materials, or reorder the content.

I am providing these lecture notes for your convenience.
Even though they look polished (typeset with LaTex), they are just notes.
There will be typos, errors, and incomplete sections.
I hope to develop them over the years to come and encourage you to read them with a critical eye.
Be critical, report potential errors, and give feed-back on form and content.

\section*{Suggested reading}
There is no one book covers the material planned for this course.
But the following books touch on the subject and are recommended:
\begin{itemize}
	\item \href{https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=1&cad=rja&uact=8&ved=2ahUKEwjd4-XS6tfkAhVI2qQKHSBoDFUQFjAAegQIAhAB&url=https%3A%2F%2Fwww.crcpress.com%2FPhysical-Biology-of-the-Cell%2FPhillips-Kondev-Theriot-Garcia%2Fp%2Fbook%2F9780815344506&usg=AOvVaw1S0mgI830cAwjb4G48vKPF}{\bf Physical Biology of the Cell} by Hernan Garcia, Jane Kondev, Julie Theriot, and Rob Phillips.
	\item \href{http://book.bionumbers.org/}{\bf Cell Biology by the Numbers} by Rob Phillips and Ron Milo.
	\item \href{https://press.princeton.edu/titles/112.html}{\bf Random Walks in Biology} by Howard Berg
\end{itemize}

% scales
\input{chapter_1}

\input{diffeq}

\input{generegulation}

% random walks and diffusion
\input{RW_diffusion}

\input{generegulation_inspace}

% liquid-liquid phase transitions
\input{liquid_liquid_phase_transitions}

% Polymers
\input{polymers}

% accuracy
\input{proofreading}

\bibliography{bib}
\end{document}



% chromatin organization
\input{chapter_4}

% gene regulation in space and time
\input{chapter_5}



